### 2024-04-03 README.Rmd
fixing readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# criticalvalue

<!-- badges: start -->
<!-- badges: end -->

The goal of criticalvalue is to ...

## Installation

You can install the development version of criticalvalue like so:

``` r
# require(devtools)
devtools::install_github("filippogambarota/criticalvalue")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
# library(criticalvalue)
devtools::load_all()

# t-test

x <- rnorm(30, 0.5, 1)
y <- rnorm(30, 0, 1)

ttest <- t.test(x, y)
critical(ttest)

# cor.test
ctest <- cor.test(x, y)
critical(ctest)

# linear model (unstandardized)
z <- rnorm(30)
q <- rnorm(30)

dat <- data.frame(x, y, z, q)
fit <- lm(y ~ x + q + z, data = dat)
fit <- critical(fit)
fit
summary(fit)

# linear model (standardized)
fit <- lm(y ~ x + q + z, data = dat)
fit <- critical(fit, standardize = TRUE)
summary(fit)
```